---
title: "CIP-stitutability and CIP purity"
output: flexdashboard::flex_dashboard
runtime: shiny
resource_files:
- data/cip_noc_processed.csv
- data/correlations.csv
---

```{r, include=FALSE}
library(tidyverse)
library(here)
library(vroom)
library(janitor)
library(plotly)
library(DescTools)
```

```{r}
cor_tbbl <- vroom(here("data","correlations.csv"))
cips <- unique(cor_tbbl$`Field of Study`)
cip_noc <- vroom(here("data","cip_noc_processed.csv"))
ginis <- cip_noc|>
  summarize(across(-NOC, Gini))|>
  pivot_longer(cols=everything(), names_to = "CIP", values_to = "Gini")|>
  filter(Gini!=1)
```


Inputs {.sidebar}
=====================================


*  Statistics Canada table 98-10-0403 reports employment counts by field of study (CIP) and occupation (NOC).

*  The goal of this exercise is to 

1) identify CIPs that are highly correlated in terms of their NOC counts, and 
2) investigate how targeted/specialized CIPs are in terms of inequality across occupations.  

On the first page, the plot on the left gives the top 10 correlated CIPs and the plot on the right shows the NOC counts for two CIPs.

On the second page, the plot shows the Gini coefficient for each CIP.


```{r}
selectInput(
  "cip",
  "Select a field of study",
  cips,
  selected = "45.06 Economics"
)
reactive_cor <- reactive({
  cor_tbbl|>
    filter(`Field of Study`==input$cip)
})

renderUI({
  selectInput("cip2", 
              "Select one of the top 10 alternatives", 
              choices = reactive_cor()$`Field of Study 2`, 
              selected= reactive_cor()$`Field of Study 2`[1])
})

reactive_point <- reactive({
  reactive_cor()|>
    filter(`Field of Study 2`==input$cip2)
})

reactive_nocs <- reactive({
  cip_noc|>
    select(NOC, input$cip, input$cip2)
})

reactive_title <- reactive({
  reactive_cor()|>
    filter(`Field of Study 2`==input$cip2)|>
    pull(Correlation)|>
    round(digits=3)
})

reactive_ginis <- reactive({
  ginis|>
    filter(CIP %in% c(input$cip, input$cip2))
})

```

Similarity between Fields of Study
=====================================  
 
Column
-------------------------------------
    
### `r renderUI({input$cip})`
    
```{r}
renderPlotly({
  plt <- ggplot2::ggplot(reactive_cor(), aes(Correlation, 
                                             fct_reorder(`Field of Study 2`, Correlation), 
                                             colour=same_broad,
                                             text=paste(sjmisc::word_wrap(`Field of Study 2`,40),
                                                        "\n Correlation:",round(Correlation, 3))
                                             ),
                         )+
    geom_point()+
    geom_point(data=reactive_point(), size=3, alpha=.5)+
    labs(title="Top 10 Spearman correlations",
       x=NULL, 
       y=NULL,
       colour="2 digit CIP")+
    theme_minimal()+ 
    theme(axis.text.y=element_blank(), 
      axis.ticks.y=element_blank()) 
  ggplotly(plt, tooltip = "text")|>
    config(displayModeBar = F)|>
    layout(legend = list(orientation = "h", x = 0.4, y = -0.2))
})
```

Column
-------------------------------------

### `r renderUI({paste("Spearman Correlation:",reactive_title())})`

```{r}
renderPlotly({
  alpha <- reactive_nocs()|>
    filter(get(input$cip)>0,
           get(input$cip2)>0
           )|>
    nrow()|>
    log()
 
  plt <- ggplot(reactive_nocs(), aes(get(input$cip), 
                                     get(input$cip2),
                                     text=paste("NOC:",
                                                sjmisc::word_wrap(NOC,40),
                                                "\n CIP:",
                                                sjmisc::word_wrap(input$cip,40),
                                                "=",
                                                get(input$cip),
                                                "\n CIP",
                                                sjmisc::word_wrap(input$cip2,40),
                                                "=",
                                                get(input$cip2)
                                     )))+
    geom_point(alpha=1/alpha)+
    labs(x=paste("NOC counts for",input$cip),
         y=paste("NOC counts for",input$cip2))+
    scale_x_continuous(trans="log10", labels = scales::comma)+
    scale_y_continuous(trans="log10", labels = scales::comma)+
    theme_minimal()
  ggplotly(plt, tooltip = "text")|>
    config(displayModeBar = F)
})
```

Field of Study Purity
=====================================================

```{r}
renderPlotly({
  plt <- ggplot(ginis, aes(Gini,
                  fct_reorder(CIP, Gini),
                  text=paste(
                    "CIP:",
                    CIP,
                    "\n Gini=",
                    round(Gini,3)
                  )
                  ))+
  geom_point(alpha=.5, size=.5)+
  geom_point(data=reactive_ginis(), colour="red", size=2)+  
  labs(x="Gini Coefficient",
       y=NULL,
       title="The Gini Coefficient measures how targeted a CIP is: if Gini=1 all grads employed in single NOC")+
    theme_minimal()+
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

plotly::ggplotly(plt, tooltip = "text")|>
  plotly::config(displayModeBar = F)
})
```




